Figures from: Vertebrate host availability and the thermal properties of mosquito-borne parasite transmission

### 2023-06-02 manuscript-figures.Rmd
clean up manuscript figures (up to trait TPC plot)

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, cache = FALSE, warning = FALSE, fig.pos = "H",
  out.extra = ""
)
# Load Libraries----------------------------------------------------------------
library(tidyverse)
library(latex2exp) #nec
library(reshape2) #nec
library(cowplot) #nec
library(viridis) #nec
library(RColorBrewer) #nec
library(MetBrewer) #nec
library(here)
library(lemon)  #nec
library(zplyr) # from https://github.com/burchill/zplyr/ # only needed for placing text using absolute coordinates in contact comparison figure

# Tell R where the manuscript lives
# here::i_am("manuscript-figures.Rmd")

# Load in necessary functions and files ----------------------------------------
# Functions for computing equilibrium quantities:
#  mosquito abundance (compute.V0), basic reproduction number (compute.R0)
# source("code/output-functions.R")

# Helper function: add a "point at infinity" for plotting when sigmaH is infinite
inf_skew_func <- function(df, skew, divide, n_rows_to_add) {
  # Get the largest value of log10(sigmaH)
  max_log10sigmaH <- filter(df, is.finite(sigmaH)) %>%
    select(sigmaH) %>%
    max() %>%
    log10()

  # Used to create a gap dividing finite and infinite sigmaH
  infinite_divide <- 10^mean(log10(c(divide, skew)))

  # Create a proxy sigmaH variable that replaces Inf with skew
  df <- mutate(df, sigmaH_proxy = ifelse(is.finite(sigmaH),
                                         sigmaH, skew
  ))

  added_sigma_vals <- seq(divide, skew, length.out = n_rows_to_add)

  for (i in 1:n_rows_to_add) {
    bump_val <- 1+0.05*(i-1)
    temp <- filter(df, is.infinite(sigmaH)) %>%
      mutate(sigmaH_proxy = ifelse(is.finite(sigmaH), sigmaH,
                                   added_sigma_vals[i]))

    df <- add_row(df, temp) %>% distinct()
  }

  return(df)

}

@@

@@

# Helper function to place legends in empty facets of plot grids
# Code by: Artem Sokolov, found here: https://stackoverflow.com/questions/54438495/shift-legend-into-empty-facets-of-a-faceted-plot-in-ggplot2
shift_legend <- function(p) {
  pnls <- cowplot::plot_to_gtable(p) %>%
    gtable::gtable_filter("panel") %>%
    with(setNames(grobs, layout$name)) %>%
    purrr::keep(~ identical(.x, zeroGrob()))

@@

```

```{r load-data, include=FALSE}
# Load results data -------------------------------------------------------------
eps <- .Machine$double.eps

# Vector trait TPC data
data.Vec <- read_rds("results/VecTPC_vals.rds")
# R0 data # !!! needs to be updated to include 1e-1, 1e0, 1e1
data.R0 <- read_rds("results/R0_vals.rds")
# Topt data
data.Topt <- read_rds("results/Topt_vals.rds")
# Critical thermal extrema data
data.CT <- read_rds("results/CT_vals.rds")

# Configure data sets for plotting ---------------------------------------------

## Get mean values of vector traits across samples
mean.Vec <- data.Vec %>% 
  select(-c(mosquito_species, pathogen, muL, etaL)) %>% 
  pivot_longer(cols = lf:V0, names_to = "variable", values_to = "value") %>%
  group_by(system_ID, Temperature, variable) %>% 
  summarise(mean = mean(value), .groups = "keep") %>% 
  ungroup() %>% 
  pivot_wider(names_from = "variable", values_from = "mean") %>% 
  mutate(etaL = rhoL / (deltaL + eps)) %>%
  # Aquatic-stage mosquito mortality rate. This should be ~infinite if deltaL = 0
  mutate(muL = etaL - rhoL) %>%
  mutate(V0 = ifelse( sigmaV_f * deltaL < (1 / lf),
                      0,
                      KL * rhoL * lf * (1 - 1 / (lf * sigmaV_f * deltaL))))

median.Vec <- data.Vec %>% 
  select(-c(mosquito_species, pathogen, muL, etaL)) %>% 
  pivot_longer(cols = lf:V0, names_to = "variable", values_to = "value") %>%
  group_by(system_ID, Temperature, variable) %>% 
  summarise(median = median(value), .groups = "keep") %>% 
  ungroup() %>% 
  pivot_wider(names_from = "variable", values_from = "median") %>% 
  mutate(etaL = rhoL / (deltaL + eps)) %>%
  # Aquatic-stage mosquito mortality rate. This should be ~infinite if deltaL = 0
  mutate(muL = etaL - rhoL) %>%
  mutate(V0 = ifelse( sigmaV_f * deltaL < (1 / lf),
                      0,
                      KL * rhoL * lf * (1 - 1 / (lf * sigmaV_f * deltaL))))

## Compute mosquito thermal tolerance range (minimum and maximum)
MosqThermRange_df <- mean.Vec %>%
  mutate(system_label = case_when(
    system_ID == "Aedes aegypti / DENV" ~ "(a) *Ae. aegypti* and DENV",
    system_ID == "Aedes aegypti / ZIKV" ~ "(b) *Ae. aegypti* and ZIKV",
    system_ID == "Aedes albopictus / DENV" ~ "(c\u200b) *Ae. albopictus* and DENV",
    system_ID == "Anopheles gambiae / Plasmodium falciparum" ~ "(d) *An. gambiae* and *P. falciparum*",
    system_ID == "Culex quinquefasciatus / WNV" ~ "(e) *Cx. quinquefasciatus* and WNV"
  )) %>%
  group_by(system_label) %>%
  filter(V0 > 0) %>%
  mutate(CTmin_V0 = min(Temperature)) %>%
  mutate(CTmax_V0 = max(Temperature)) %>%
  select(system_ID, system_label, CTmin_V0, CTmax_V0) %>%
  distinct()

# Title names for each mosquito+pathogen combination (NB: to be added to get-analysis-dfs.R)
data.R0 <- data.R0 %>%
  ungroup() %>% 
  mutate(system_label = case_when(
    system_ID == "Aedes aegypti / DENV" ~ "(a) *Ae. aegypti* and DENV",
    system_ID == "Aedes aegypti / ZIKV" ~ "(b) *Ae. aegypti* and ZIKV",
    system_ID == "Aedes albopictus / DENV" ~ "(c\u200b) *Ae. albopictus* and DENV",
    system_ID == "Anopheles gambiae / Plasmodium falciparum" ~ "(d) *An. gambiae* and *P. falciparum*",
    system_ID == "Culex quinquefasciatus / WNV" ~ "(e) *Cx. quinquefasciatus* and WNV"
  ))

# ## Add a "point at infinity" for sigmaH
# # Value of sigmaH to replace 'Inf' with for plotting
# infinite_skew <- 10^3 + 2500
# 
# # Get the largest value of log10(sigmaH)
# max_log10sigmaH <- filter(AllOutputs, is.finite(sigmaH)) %>%
#   select(sigmaH) %>%
#   max() %>%
#   log10()
# 
# # Used to create a gap dividing finite and infinite sigmaH
# infinite_divide <- 10^mean(c(3, log10(infinite_skew)))
# 
# # add a second proxy sigmaH point at infinity to make the values easier to see there
# temp <- filter(AllOutputs, is.infinite(sigmaH)) %>%
#   mutate(sigmaH_proxy = ifelse(is.finite(sigmaH), sigmaH, 1.2 * infinite_divide))
# # Create a proxy sigmaH variable that replaces Inf with infinite_skew
# AllOutputs <- mutate(AllOutputs, sigmaH_proxy = ifelse(is.finite(sigmaH),
#                                                        sigmaH, infinite_skew
# ))
# AllOutputs <- add_row(AllOutputs, temp) %>% distinct()
# 
# # Define sigmaH tick breaks and labels
# sigmaH_breaks <- c(10^c(seq(0, 3, by = 1)), infinite_divide, infinite_skew)
# sigmaH_labels <- c(paste(c(10^c(seq(0, 3, by = 1)))), TeX("$\\ldots$"), TeX("$\\infty$"))
# 
# ## Parasite thermal characteristics
# ThermPathChars <- read_csv("results/ThermalCharacteristics.csv", show_col_types = FALSE)
# 
# # Title names for each mosquito+pathogen combination
# ThermPathChars <- ThermPathChars %>%
#   mutate(system_label = case_when(
#     system_ID == "Aedes aegypti+DENV" ~ "(a) *Ae. aegypti* and DENV",
#     system_ID == "Aedes aegypti+ZIKV" ~ "(b) *Ae. aegypti* and ZIKV",
#     system_ID == "Aedes albopictus+DENV" ~ "(c\u200b) *Ae. albopictus* and DENV",
#     system_ID == "Anopheles spp.+P.fal." ~ "(d) *An. gambiae* and *P. falciparum*",
#     system_ID == "Culex quinquefasciatus+WNV" ~ "(e) *Cx. quinquefasciatus* and WNV"
#   ))
# ## Add a "point at infinity" for sigmaH (NB: to be added to get-analysis-dfs.R)
# # Value of sigmaH to replace 'Inf' with for plotting
# infinite_skew <- 10^3 + 2500
# 
# # Get the largest value of log10(sigmaH)
# max_log10sigmaH <- filter(ThermPathChars, is.finite(sigmaH)) %>%
#   select(sigmaH) %>%
#   max() %>%
#   log10()
# 
# # Used to create a gap dividing finite and infinite sigmaH
# infinite_divide <- 10^mean(c(3, log10(infinite_skew)))
# 
# # Create a proxy sigmaH variable that replaces Inf with infinite_skew
# ThermPathChars <- mutate(ThermPathChars, sigmaH_proxy = ifelse(is.finite(sigmaH),
#                                                                sigmaH, infinite_skew
# ))
# # add a second proxy sigmaH point at infinity to make the values easier to see there
# temp <- filter(ThermPathChars, is.infinite(sigmaH)) %>%
#   mutate(sigmaH_proxy = ifelse(is.finite(sigmaH), sigmaH, 1.2 * infinite_divide))
# 
# ThermPathChars <- add_row(ThermPathChars, temp) %>%
#   arrange(sigmaH_proxy) %>%
#   distinct()
# 
# ## Options
# # Define sigmaH tick breaks and labels
# sigmaH_breaks <- c(10^c(seq(-1, 3, by = 1)), infinite_divide, infinite_skew)
# sigmaH_labels <- c(paste(c(10^c(seq(-1, 3, by = 1)))), TeX("$\\ldots$"), TeX("$\\infty$"))

```

@@

```{r fig-1_contact_comparison, dev = c('pdf','svg', 'png'), out.width='100%', fig.align='center', fig.height=5, fig.width=7, fig.align='center', fig.cap='\\label{fig:contact_comparison} A comparison of the Ross-Macdonald and Chitnis contact rate formulations.'}

# resolution setting
resolution <- 1E2

## Define different contact models
# For each model, there is a vector-to-host and host-to-vector term

@@

# x = Host population density, y = Bites per mosquito per day
plot_HD_MC <- new_df %>%
  filter(x_vars == "Host population density") %>%
  filter(y_vars == "Bites per mosquito per day") %>%
  ggplot(aes(x = x_vals, y = y_vals, colour = `Transmission model`)) +

  # curves:
  geom_path(size = 1) +
  geom_blank(aes(y = y_max)) +

  # plot labels
  geom_abs_text(aes(xpos = 0.08, ypos = 0.89, label = label),
                color = "black", show.legend = FALSE
  ) +

  # x-axis:
  scale_x_continuous(
    name = "",
    expand = expansion(mult = c(0, .025))
  ) +
  # y-axis:
  scale_y_continuous(
    name = "Bites per mosquito per day",
    expand = c(0, 0),
    limits = c(0, NA)
  )

# x = Mosquito population density, y = Bites per mosquito per day
plot_MD_MC <- new_df %>%
  filter(x_vars == "Mosquito population density") %>%
  filter(y_vars == "Bites per mosquito per day") %>%
  ggplot(aes(x = x_vals, y = y_vals, colour = `Transmission model`)) +

  # curves:
  geom_path(size = 1) +
  geom_blank(aes(y = y_max)) +

  # plot labels
  geom_abs_text(aes(xpos = 0.08, ypos = 0.89, label = label),
                color = "black", show.legend = FALSE
  ) +

  # x-axis:
  scale_x_continuous(
    name = "",
    expand = expansion(mult = c(0, .025))
  ) +
  # y-axis:
  scale_y_continuous(
    name = "",
    expand = c(0, 0),
    limits = c(0, NA)
  ) +
  # hide legend
  guides(color = FALSE)

# x = Mosquito population density, y = Bites per host per day
plot_MD_HC <- new_df %>%
  filter(x_vars == "Mosquito population density") %>%
  filter(y_vars == "Bites per host per day") %>%
  # x = density variables, y = contact rates
  ggplot(aes(x = x_vals, y = y_vals, colour = `Transmission model`)) +

  # curves:
  geom_path(size = 1) +
  geom_blank(aes(y = y_max)) +

  # plot labels
  geom_abs_text(aes(xpos = 0.08, ypos = 0.89, label = label),
                color = "black", show.legend = FALSE
  ) +

  # x-axis:
  scale_x_continuous(
    name = "Mosquito population density",
    expand = expansion(mult = c(0, .025))
  ) +
  # y-axis:
  scale_y_continuous(
    name = "",
    expand = c(0, 0),
    limits = c(0, NA)
  )

# x = Host population density, y = Bites per host per day
plot_HD_HC <- new_df %>%
  filter(x_vars == "Host population density") %>%
  filter(y_vars == "Bites per host per day") %>%
  # x = density variables, y = contact rates
  ggplot(aes(x = x_vals, y = y_vals, colour = `Transmission model`)) +

  # curves:
  geom_path(size = 1) +
  geom_blank(aes(y = y_max)) +

  # plot labels
  geom_abs_text(aes(xpos = 0.08, ypos = 0.89, label = label),
                color = "black", show.legend = FALSE
  ) +

  # x-axis:
  scale_x_continuous(
    name = "Host population density",
    expand = expansion(mult = c(0, .025))
  ) +
  # y-axis:
  scale_y_continuous(
    name = "Bites per host per day",
    expand = c(0, 0),
    limits = c(0, NA)
  ) +
  # hide legend
  guides(color = FALSE)

contact_plot <- plot_grid(
  plot_HD_MC + theme(legend.position = "none"),
  plot_MD_MC + theme(legend.position = "none"),
  plot_HD_HC + theme(legend.position = "none"),
  plot_MD_HC + theme(legend.position = "none")
)

legend <- get_legend(
  plot_HD_MC +
    guides(color = guide_legend(nrow = 1)) +
    theme(legend.position = "bottom")
)

contact_plot <- plot_grid(contact_plot, legend, ncol = 1, rel_heights = c(1, .05))

contact_plot
```

```{r fig-3_MosquitoThermalTraits, dev = c('pdf','svg', 'png'),  out.width='100%', fig.height=6, fig.width=9, fig.align='center', fig.cap='\\label{fig:MosquitoThermalTraits} Thermal performance curves for the mosquito and parasite traits used in our model. '}

## Set up plot dataframe
MosqThermTraits_df <- median.Vec %>%
  ## Set up plot dataframe
  # Mortality rate
  mutate(muV = 1/lf) %>% 
  # Recruitment rate
  mutate(lambdaV = V0 * muV) %>%
  # Prob. of surviving EIP
  mutate(P_EIP = exp(-muV / etaV)) %>%
  # Select relevant traits and parameters
  select(
    Temperature, system_ID, lambdaV, sigmaV, betaV, lf, P_EIP
  ) %>%
  # Turn all columns except Temperature and system_ID (Mosquito sp. + parasite sp.) into
  # 'variables'
  melt(id = c("Temperature", "system_ID")) %>%
  group_by(Temperature, system_ID)

# Assign nice labels for each variable
levels(MosqThermTraits_df$variable) <- c(
  lambdaV = TeX("(a) Recruitment rate: $\\lambda_V$ (ind. day$^{-1}$)"),
  sigmaV = TeX("(b) Maximum biting frequency: $\\sigma_V$ (day$^{-1}$)"),
  betaV = TeX("(c) Vector competence: $\\beta_V$"),
  lf = TeX("(d) Average adult lifespan: $1/\\mu_V$ (days)"),
  P_EIP = TeX("(e) Prob. of surviving EIP: $\\theta_V$")
)

MosqThermTraits_plot <- MosqThermTraits_df %>%
  # Set up plotting variables and theme
  # x = Temperature, y = trait value
  ggplot(mapping = aes(x = Temperature, y = value, group = system_ID)) +

  # Plot trait values:
  geom_line(aes(colour = system_ID), size = 1) +

  # x-axis:
  scale_x_continuous(
    name = "Temperature (°C)",
    expand = c(0, 0),
    limits = c(10, 35),
    breaks = seq(10, 36, 2)
  ) +

  # y-axis:
  scale_y_continuous(expand = c(0.05, 0)) +

  # color:
  scale_color_manual(
    name = "System:",
    values = met.brewer("Egypt", 5, direction = -1),
    labels = function(x) {
      case_when(
        x == "Aedes aegypti / DENV" ~ "*Aedes aegypti* and DENV",
        x == "Aedes aegypti / ZIKV" ~ "*Ae. aegypti* and ZIKV",
        x == "Aedes albopictus / DENV" ~ "*Aedes albopictus* and DENV",
        x == "Anopheles gambiae / Plasmodium falciparum" ~ "*Anopheles gambiae* and *Plasmodium falciparum*",
        x == "Culex quinquefasciatus / WNV" ~ "*Culex quinquefasciatus* and WNV"
      )
    }
  ) +

  # faceting:
  facet_wrap(~variable, nrow = 3, scales = "free", labeller = labeller(variable = label_parsed)) +

  # legend:
  guides(color = guide_legend(override.aes = list(size = 4))) +

  # theme options:
  theme_minimal_vgrid(11) +
  theme(
    axis.title.y = element_blank(),
    legend.position = "bottom",
    legend.margin = margin(),
    legend.text = ggtext::element_markdown(),
    legend.direction = "vertical",
    legend.justification = "center",
    axis.title.x = element_text(hjust = 0.2),
    strip.text = element_text(
      face = "bold",
      hjust = 0
    )
  )

@@

```{r fig-4_R0_TPCs, dev = c('pdf','svg', 'png'),  message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:R0_plots} ${R}_0$ as a function of temperature has the properties of a thermal performance curve (TPC): it is positive on a single bounded interval and has a unique maximum value.'}

# Plot R0 vs. temperature curves. One plot for each mosquito+pathogen pair.
# One curve for Ross-Macdonald, a few for Chitnis

@@

# Plot
R0_plot <- shift_legend(R0_plot)

```

@@

# add many proxy sigmaH points at infinity to make the values easier to see there
max_sigmaH <- max(filter(Toptalt_df, is.finite(sigmaH))$sigmaH)
# Value of sigmaH to replace 'Inf' with for plotting
infinite_skew <- 10^2.35#10^log10(max_sigmaH*1.5) # 1.1*max_sigmaH #10^2 + 80
# Used to create a gap dividing finite and infinite sigmaH
infinite_divide <- 10^2.7 #10^log10(max_sigmaH*1)

@@

min_Topt <- floor(min(min(Toptalt_df$mean),min(Toptalt_df$median)))
max_Topt <- ceiling(max(max(Toptalt_df$mean),max(Toptalt_df$median)))
Topt_breaks <- seq(min_Topt, max_Topt, by = 1)

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, cache = FALSE, warning = FALSE, fig.pos = "H",
  out.extra = ""
)
# Load Libraries----------------------------------------------------------------
library(tidyverse)
library(latex2exp)
library(reshape2)
library(cowplot)
library(viridis)
library(RColorBrewer)
library(MetBrewer)
library(here)
library(lemon)
library(zplyr) # from https://github.com/burchill/zplyr/ # only needed for placing text using absolute coordinates in contact comparison figure

@@

# !!! [] make sure all plots have font size set at 11 max, unless necessary

@@

# Helper function: adds a "point at infinity" for plots with infinite sigmaH
inf_skew_func <- function(df, skew, divide, n_rows_to_add) {
  # Get the largest value of log10(sigmaH) in the data frame
  max_log10sigmaH <- filter(df, is.finite(sigmaH)) %>%
    select(sigmaH) %>%
    max() %>%
    log10()

  # Define the gap dividing finite and infinite sigmaH
  infinite_divide <- 10^mean(log10(c(divide, skew)))

  # Create a proxy sigmaH variable that replaces Inf with the value of "skew"
  df <- mutate(df, sigmaH_proxy = ifelse(is.finite(sigmaH),
    sigmaH, skew
  ))

  # add more proxy sigmaH values to make a "spread" at Inf
  added_sigma_vals <- seq(divide, skew, length.out = n_rows_to_add)
  for (i in 1:n_rows_to_add) {
    bump_val <- 1 + 0.05 * (i - 1)
    temp <- filter(df, is.infinite(sigmaH)) %>%
      mutate(sigmaH_proxy = ifelse(is.finite(sigmaH), sigmaH,
        added_sigma_vals[i]
      ))
    df <- add_row(df, temp) %>% distinct()
  }
  return(df)
}

# Helper function: place legends in empty facets of plot grids
# Code by: Artem Sokolov, found here: https://stackoverflow.com/questions/54438495/shift-legend-into-empty-facets-of-a-faceted-plot-in-ggplot2
shift_legend <- function(p) {
  pnls <- cowplot::plot_to_gtable(p) %>%
    gtable::gtable_filter("panel") %>%
    with(setNames(grobs, layout$name)) %>%
    purrr::keep(~ identical(.x, zeroGrob()))

@@

# Helper function: apply nice labels for pathosystems
pathosystem_labeller <- function(df) {
  df <- df %>%
    ungroup() %>%
    mutate(system_label = case_when(
      system_ID == "Aedes aegypti / DENV" ~ "(a) *Ae. aegypti* and DENV",
      system_ID == "Aedes aegypti / ZIKV" ~ "(b) *Ae. aegypti* and ZIKV",
      system_ID == "Aedes albopictus / DENV" ~ "(c&#41; *Ae. albopictus* and DENV",
      system_ID == "Anopheles gambiae / Plasmodium falciparum" ~ "(d) *An. gambiae* and *P. falciparum*",
      system_ID == "Culex quinquefasciatus / WNV" ~ "(e) *Cx. quinquefasciatus* and WNV"
    ))
  return(df)
}

# Load required data ----

# !!! [] make sure all data is getting loaded here

eps <- .Machine$double.eps

# Vector trait TPC data
data.Vec <- read_rds("results/VecTPC_vals.rds") %>%
  #
  pathosystem_labeller()
# R0 data
# !!! [] needs to be updated to include 1e-1, 1e0, 1e1
data.R0 <- read_rds("results/R0_vals.rds") %>%
  pathosystem_labeller()
# Topt data
data.Topt <- read_rds("results/Topt_vals.rds") %>%
  pathosystem_labeller()
# Critical thermal extrema data
data.CT <- read_rds("results/CT_vals.rds") %>%
  pathosystem_labeller()

# Configure data sets for plotting ----

## Get values of vector traits across samples
# Mean values
mean.Vec <- data.Vec %>%
  select(-c(mosquito_species, pathogen, muL, etaL)) %>%
  pivot_longer(cols = lf:V0, names_to = "variable", values_to = "value") %>%
  group_by(system_ID, Temperature, variable) %>%
  summarise(mean = mean(value), .groups = "keep") %>%
  ungroup() %>%
  pivot_wider(names_from = "variable", values_from = "mean") %>%
  mutate(etaL = rhoL / (deltaL + eps)) %>%
  mutate(muL = etaL - rhoL) %>%
  mutate(V0 = ifelse(sigmaV_f * deltaL < (1 / lf),
    0,
    KL * rhoL * lf * (1 - 1 / (lf * sigmaV_f * deltaL))
  ))
# Median values (this one is used more frequently)
median.Vec <- data.Vec %>%
  select(-c(mosquito_species, pathogen, muL, etaL)) %>%
  pivot_longer(cols = lf:V0, names_to = "variable", values_to = "value") %>%
  group_by(system_ID, Temperature, variable) %>%
  summarise(median = median(value), .groups = "keep") %>%
  ungroup() %>%
  pivot_wider(names_from = "variable", values_from = "median") %>%
  mutate(etaL = rhoL / (deltaL + eps)) %>%
  # Aquatic-stage mosquito mortality rate. This should be ~infinite if deltaL = 0
  mutate(muL = etaL - rhoL) %>%
  mutate(V0 = ifelse(sigmaV_f * deltaL < (1 / lf),
    0,
    KL * rhoL * lf * (1 - 1 / (lf * sigmaV_f * deltaL))
  ))

# Compute *mosquito* thermal tolerance range (minimum and maximum)
MosqThermRange_df <- median.Vec %>%
  pathosystem_labeller() %>%
  group_by(system_label) %>%
  filter(V0 > 0) %>%
  mutate(CTmin_V0 = min(Temperature)) %>%
  mutate(CTmax_V0 = max(Temperature)) %>%
  select(system_ID, system_label, CTmin_V0, CTmax_V0) %>%
  distinct()
```

@@

```{r fig-1_contact_comparison, dev = c('pdf','svg', 'png'), out.width='100%', fig.align='center', fig.height=5, fig.width=7, fig.align='center', fig.cap='\\label{fig:contact_comparison} A comparison of the Ross-Macdonald and Chitnis contact rate formulations.'}
# resolution setting
resolution <- 1E2

## Define the contact models
# For each model, there is a vector-to-host and host-to-vector term

@@

# x = Host population density, y = Bites per mosquito per day
plot_HD_MC <- new_df %>%
  filter(x_vars == "Host population density") %>%
  filter(y_vars == "Bites per mosquito per day") %>%
  ggplot(aes(x = x_vals, y = y_vals, colour = `Transmission model`)) +
  # curves:
  geom_path(size = 1) +
  geom_blank(aes(y = y_max)) +
  # plot labels
  geom_abs_text(aes(xpos = 0.08, ypos = 0.89, label = label),
    color = "black", show.legend = FALSE
  ) +
  # x-axis:
  scale_x_continuous(
    name = "",
    expand = expansion(mult = c(0, .025))
  ) +
  # y-axis:
  scale_y_continuous(
    name = "Bites per mosquito per day",
    expand = c(0, 0),
    limits = c(0, NA)
  )

# x = Mosquito population density, y = Bites per mosquito per day
plot_MD_MC <- new_df %>%
  filter(x_vars == "Mosquito population density") %>%
  filter(y_vars == "Bites per mosquito per day") %>%
  ggplot(aes(x = x_vals, y = y_vals, colour = `Transmission model`)) +
  # curves:
  geom_path(size = 1) +
  geom_blank(aes(y = y_max)) +
  # plot labels
  geom_abs_text(aes(xpos = 0.08, ypos = 0.89, label = label),
    color = "black", show.legend = FALSE
  ) +
  # x-axis:
  scale_x_continuous(
    name = "",
    expand = expansion(mult = c(0, .025))
  ) +
  # y-axis:
  scale_y_continuous(
    name = "",
    expand = c(0, 0),
    limits = c(0, NA)
  ) +
  # hide legend
  guides(color = FALSE)

# x = Mosquito population density, y = Bites per host per day
plot_MD_HC <- new_df %>%
  filter(x_vars == "Mosquito population density") %>%
  filter(y_vars == "Bites per host per day") %>%
  # x = density variables, y = contact rates
  ggplot(aes(x = x_vals, y = y_vals, colour = `Transmission model`)) +
  # curves:
  geom_path(size = 1) +
  geom_blank(aes(y = y_max)) +
  # plot labels
  geom_abs_text(aes(xpos = 0.08, ypos = 0.89, label = label),
    color = "black", show.legend = FALSE
  ) +
  # x-axis:
  scale_x_continuous(
    name = "Mosquito population density",
    expand = expansion(mult = c(0, .025))
  ) +
  # y-axis:
  scale_y_continuous(
    name = "",
    expand = c(0, 0),
    limits = c(0, NA)
  )

# x = Host population density, y = Bites per host per day
plot_HD_HC <- new_df %>%
  filter(x_vars == "Host population density") %>%
  filter(y_vars == "Bites per host per day") %>%
  # x = density variables, y = contact rates
  ggplot(aes(x = x_vals, y = y_vals, colour = `Transmission model`)) +
  # curves:
  geom_path(size = 1) +
  geom_blank(aes(y = y_max)) +
  # plot labels
  geom_abs_text(aes(xpos = 0.08, ypos = 0.89, label = label),
    color = "black", show.legend = FALSE
  ) +
  # x-axis:
  scale_x_continuous(
    name = "Host population density",
    expand = expansion(mult = c(0, .025))
  ) +
  # y-axis:
  scale_y_continuous(
    name = "Bites per host per day",
    expand = c(0, 0),
    limits = c(0, NA)
  ) +
  # hide legend
  guides(color = FALSE)

# Combine plots into a grid and remove legends
contact_plot <- plot_grid(
  plot_HD_MC + theme(legend.position = "none"),
  plot_MD_MC + theme(legend.position = "none"),
  plot_HD_HC + theme(legend.position = "none"),
  plot_MD_HC + theme(legend.position = "none")
)

# Set legend to bottom of grid
legend <- get_legend(
  plot_HD_MC +
    guides(color = guide_legend(nrow = 1)) +
    theme(legend.position = "bottom")
)

# Put together grid and legend
contact_plot <- plot_grid(contact_plot, legend, ncol = 1, rel_heights = c(1, .05))

# Plot
contact_plot
```

```{r fig-3_MosquitoThermalTraits, dev = c('pdf','svg', 'png'),  out.width='100%', fig.height=6, fig.width=9, fig.align='center', fig.cap='\\label{fig:MosquitoThermalTraits} Thermal performance curves for the mosquito and parasite traits used in our model. The range of temperatures is restricted to values where the immature development rate is positive. For traits that do not involve a parasite, *Aedes aegypti* with DENV and *Ae. aegypti* with ZIKV curves overlap.'}
## Set up plot dataframe
MosqThermTraits_df <- median.Vec %>%
  # Mortality rate
  mutate(muV = 1 / lf) %>%
  # Recruitment rate
  mutate(lambdaV = V0 * muV) %>%
  # Prob. of surviving EIP
  mutate(P_EIP = exp(-muV / etaV)) %>%
  # Prob. of larval survival
  # Select relevant traits and parameters
  select(Temperature, system_ID, sigmaV, sigmaV_f, rhoL, lf, deltaL, P_EIP, betaV) %>%
  # select(Temperature, system_ID, lambdaV, sigmaV, betaV, lf, P_EIP) %>%
  # Turn all columns except Temperature and system_ID into variables
  melt(id = c("Temperature", "system_ID")) %>%
  mutate(variable = factor(variable,
    levels = c("lf", "sigmaV", "sigmaV_f", "rhoL", "deltaL", "P_EIP", "betaV"),
    labels = c(
      TeX("(a) Adult lifespan: $1/\\mu_V$ (days)"),
      TeX("(b) Biting frequency: $\\sigma_V$ (day$^{-1}$)"),
      TeX("(c) Eggs per female per day: $\\sigma_V f$"),
      TeX("(d) Immature development rate: $\\rho_L$"),
      TeX("(e) Prob. of immature survival: $\\delta_L$"),
      TeX("(f) Prob. of surviving EIP: $\\theta_V$"),
      TeX("(g) Vector competence: $\\beta_V$")
    )
  )) %>%
  group_by(Temperature, system_ID)

# !!! [] be sure to use a consistent order for mosquito life history traits
# !!! [] be sure to use a consistent list of mosquito life history traits:
#         - adult lifespan (1/mu_V)
#         - biting frequency (sigma_V)
#         - eggs per female per day (sigma_V * f)
#         - immature development rate
#         - Pr(larval survival) (deltaL)
#         - Pr(surviving EIP) (theta_V)
#         - vector competence (beta_V)

MosqThermTraits_plot <- MosqThermTraits_df %>%
  # Set up plotting variables and theme
  # x = Temperature, y = trait value
  ggplot(mapping = aes(x = Temperature, y = value, group = system_ID)) +
  # Plot trait values:
  geom_line(aes(colour = system_ID), size = 1) +
  # x-axis:
  scale_x_continuous(
    name = "Temperature (°C)",
    expand = c(0, 0),
    limits = c(10, 35),
    breaks = seq(10, 36, 2)
  ) +
  # y-axis:
  scale_y_continuous(expand = c(0.05, 0)) +
  # color:
  scale_color_manual(
    name = "System:",
    values = met.brewer("Egypt", 5, direction = -1),
    labels = function(x) {
      case_when(
        x == "Aedes aegypti / DENV" ~ "*Aedes aegypti* and DENV",
        x == "Aedes aegypti / ZIKV" ~ "*Ae. aegypti* and ZIKV",
        x == "Aedes albopictus / DENV" ~ "*Aedes albopictus* and DENV",
        x == "Anopheles gambiae / Plasmodium falciparum" ~ "*Anopheles gambiae* and *Plasmodium falciparum*",
        x == "Culex quinquefasciatus / WNV" ~ "*Culex quinquefasciatus* and WNV"
      )
    }
  ) +
  # faceting:
  facet_wrap(~variable,
    nrow = 3, scales = "free",
    labeller = labeller(variable = label_parsed)
  ) +
  # legend:
  guides(color = guide_legend(override.aes = list(size = 4))) +
  # theme options:
  theme_minimal_vgrid(11) +
  theme(
    axis.title.y = element_blank(),
    legend.position = "bottom",
    legend.margin = margin(),
    legend.text = ggtext::element_markdown(),
    legend.direction = "vertical",
    legend.justification = "center",
    axis.title.x = element_text(hjust = 0.2),
    strip.text = element_text(
      face = "bold",
      hjust = 0
    )
  )

@@

```{r fig-4_R0_TPCs, dev = c('pdf','svg', 'png'),  message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:R0_plots} ${R}_0$ as a function of temperature has the properties of a thermal performance curve (TPC): it is positive on a single bounded interval and has a unique maximum value.'}
# Plot R0 vs. temperature curves. One plot for each mosquito+pathogen pair.
# One curve for Ross-Macdonald, a few for Chitnis

@@

# Plot
R0_plot <- shift_legend(R0_plot)
```

@@

# add many proxy sigmaH points at infinity to make the values easier to see there
max_sigmaH <- max(filter(Toptalt_df, is.finite(sigmaH))$sigmaH)
# Value of sigmaH to replace 'Inf' with for plotting
infinite_skew <- 10^2.35 # 10^log10(max_sigmaH*1.5) # 1.1*max_sigmaH #10^2 + 80
# Used to create a gap dividing finite and infinite sigmaH
infinite_divide <- 10^2.7 # 10^log10(max_sigmaH*1)

@@

min_Topt <- floor(min(min(Toptalt_df$mean), min(Toptalt_df$median)))
max_Topt <- ceiling(max(max(Toptalt_df$mean), max(Toptalt_df$median)))
Topt_breaks <- seq(min_Topt, max_Topt, by = 1)
